---
title: "IMM+ RCT impact analysis - stat_summit_discussion - effect size"
output:
  html_document:
    df_print: paged
date: "2023-08-18"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 11)
knitr::opts_knit$set(root.dir = here::here())
```

## Set up

- Run CONTROL.r


```{r load-packages}
pacman::p_load(
  "tidyverse",
  "qualtr",
  "salesforcer",
  "glue",
  "tidytext",
  "lubridate",
  "pander",
  "infer",
  "effsize"
)

register_options()

library(ggdist)
library(distributional)

# SOURCE SCRIPTS ----------------------------------------------------------

source("Z:/FSC Pilot - CCDF/Analysis/rct-analysis/R/pipeline/utils.R")
source("Z:/FSC Pilot - CCDF/Analysis/rct-analysis/R/pipeline/load-data.R")
source("Z:/FSC Pilot - CCDF/Analysis/rct-analysis/R/pipeline/clean-data.R")

source("Z:/FSC Pilot - CCDF/Analysis/rct-analysis/R/pipeline/rct_prep.R")

# RUN PIPELINE ------------------------------------------------------------

dat_raw <- load_data(use_cache = TRUE) # sometimes first time running script returns an error for some reason that I cannot pinpoint....
# try running twice

dat_clean <- clean_data(dat_raw, use_cache = FALSE)

dat_rct_prep <- rct_prep(dat_clean, use_cache = FALSE)
```

## Average change

- LEFT: Average change measured as a percentage of the size of the scale's range to let us discuss the change in terms of the scale's definition and in its unit
- RIGHT: Average change as a proportion of the scale range to let us compare changes across outcomes 

```{r tibble}
d_pre_emp <- tibble::tribble(
  ~outcome, ~coefficient, ~estimate, ~std_error, ~t, ~p,
  "Self Esteem", "Intercept", .6301, .4509, 1.397, .164,
  "Self Esteem", "Treatment", 1.1934, .6148, 1.941, .054,
  "Resilience", "Intercept", .01716, .08613, .199, .842,
  "Resilience", "Treatment", .15348, .11587, 1.325, .1873,
  "Emotional Intelligence", "Intercept", -.1618, .4821, -.336, .7377,
  "Emotional Intelligence", "Treatment", 1.6794, .6468, 2.596, .0103,
  "Locus of Control", "Intercept", 0.1549, 0.5726, 0.271, 0.787,
  "Locus of Control", "Treatment", 0.4716, 0.7800, 0.605, 0.546,
  "Healthy Behaviour", "Intercept", 0.05797, 0.27634, 0.210, 0.834,
  "Healthy Behaviour", "Treatment", 0.28686, 0.37003, 0.775, 0.439,
  "Mental Health", "Intercept", -0.3714, 0.3870, -0.960, 0.339,
  "Mental Health", "Treatment", -0.7833, 0.5240, -1.495, 0.137
)

d_ehs <- tibble::tribble(
  ~outcome, ~coefficient, ~estimate, ~std_error, ~t, ~p,
  "Total EHS", "Intercept", -0.2498, 0.2115, -1.181, 0.24,
  "Total EHS", "Treatment", 1.3409, 0.2895, 4.632, 7.95e-06,
  "Empowerment", "Intercept", -0.2494, 0.2111, -1.182, 0.23926,
  "Empowerment", "Treatment", 0.7826, 0.2889, 2.709, 0.00755,
  "Motivation", "Intercept", -0.5616, 0.2945, -1.907, 0.058510,
  "Motivation", "Treatment", 1.4113, 0.4031, 3.501, 0.000615,
  "Utilization", "Intercept", -0.1105, 0.2761, 0.400, 0.689600,
  "Utilization", "Treatment", 1.4829, 0.3780, 3.923, 0.000134,
  "Goal Orientation", "Intercept", -0.2023, 0.2869, -0.705, 0.482,
  "Goal Orientation", "Treatment", 1.7150, 0.3926, 4.368, 2.37e-05
)
```

```{r average-graph}
# pre-employability skills

dplt_01 <- d_pre_emp %>%
  select(outcome, coefficient, estimate) %>%
  pivot_wider(names_from = coefficient, values_from = estimate) %>%
  rename(
    treatment_effect = `Treatment`,
    control = `Intercept`
  ) %>%
  mutate(
    treatment = treatment_effect + control,
    scale_range = case_when(
      outcome == "Self Esteem" ~ 30 - 5, # lowest end of scale - 1 so range covers the full scale
      outcome == "Resilience" ~ 5 - 0,
      outcome == "Emotional Intelligence" ~ 45 - 8,
      outcome == "Locus of Control" ~ 45 - 8,
      outcome == "Healthy Behaviour" ~ 15 - 2,
      outcome == "Mental Health" ~ 12 - -1
    )
  ) %>%
  pivot_longer(
    cols = c(control, treatment),
    names_to = "group",
    values_to = "Scale units"
  ) %>%
  mutate(
    `% of scale range` = `Scale units` / scale_range * 100
  ) %>%
  pivot_longer(
    cols = c(`Scale units`, `% of scale range`),
    names_to = "transformation",
    values_to = "observed_change"
  ) %>%
  mutate(
    transformation = fct_inorder(transformation)
  )

dplt_01 %>%
  ggplot(aes(y = group, x = observed_change)) +
  facet_grid(rows = vars(outcome), cols = vars(transformation), scales = "free_x", switch = "y") +
  geom_segment(aes(xend = 0, yend = after_stat(y))) +
  geom_point(shape = 21, size = 3, fill = "#0000FF") +
  geom_text(
    aes(label = round(observed_change, 1)),
    position = position_nudge(y = .12)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(.1, .1))
  ) +
  labs(
    x = "Average Change in Score",
    y = NULL
  ) +
  bptheme::theme_blueprint(grid = "Xxy", grid_col = "grey80") +
  theme(strip.placement = "outside")

# EHS

dplt_02 <- d_ehs %>%
  select(outcome, coefficient, estimate) %>%
  pivot_wider(names_from = coefficient, values_from = estimate) %>%
  rename(
    treatment_effect = `Treatment`,
    control = `Intercept`
  ) %>%
  mutate(
    treatment = treatment_effect + control,
    scale_range = case_when(
      outcome == "Total EHS" ~ 10 - -1,
      outcome == "Empowerment" ~ 10 - -1,
      outcome == "Motivation" ~ 10 - -1,
      outcome == "Utilization" ~ 10 - -1,
      outcome == "Goal Orientation" ~ 10 - -1
    )
  ) %>%
  pivot_longer(
    cols = c(control, treatment),
    names_to = "group",
    values_to = "EHS Scale units"
  ) %>%
  mutate(
    `% of scale range` = `EHS Scale units` / scale_range * 100
  ) %>%
  pivot_longer(
    cols = c(`EHS Scale units`, `% of scale range`),
    names_to = "transformation",
    values_to = "observed_change"
  ) %>%
  mutate(
    transformation = fct_inorder(transformation)
  )

dplt_02 %>%
  ggplot(aes(y = group, x = observed_change)) +
  facet_grid(rows = vars(outcome), cols = vars(transformation), scales = "free_x", switch = "y") +
  geom_segment(aes(xend = 0, yend = after_stat(y))) +
  geom_point(shape = 21, size = 3, fill = "#0000FF") +
  geom_text(
    aes(label = round(observed_change, 1)),
    position = position_nudge(y = .12)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(.1, .1))
  ) +
  labs(
    x = "Average Change in Score",
    y = NULL
  ) +
  bptheme::theme_blueprint(grid = "Xxy", grid_col = "grey80") +
  theme(strip.placement = "outside")
```

# What does it mean about CCDF's model in general?

```{r distribution}
dplt_03 <- d_pre_emp %>%
  mutate(
    scale_range = case_when(
      outcome == "Self Esteem" ~ 30 - 5,
      outcome == "Resilience" ~ 5 - 0,
      outcome == "Emotional Intelligence" ~ 45 - 8,
      outcome == "Locus of Control" ~ 45 - 8,
      outcome == "Healthy Behaviour" ~ 15 - 2,
      outcome == "Mental Health" ~ 12 - -1
    ),
    estimate__range_pct = estimate / scale_range * 100,
    std_error__range_pct = std_error / scale_range * 100
  ) |>
  rename(
    estimate__scale_units = estimate,
    std_error__scale_units = std_error
  ) |>
  pivot_longer(
    cols = matches("estimate|std_error"),
    names_to = c(".value", "transformation"),
    names_pattern = "(.*)__(.*)"
  ) |>
  mutate(
    transformation = fct_inorder(transformation)
  )

dplt_03 %>%
  ggplot(aes(y = coefficient)) +
  facet_grid(rows = vars(outcome), cols = vars(transformation), switch = "y", scales = "free_x") +
  stat_halfeye(
    aes(
      xdist = dist_normal(estimate, std_error),
      fill = after_stat(x > 0)
    ),
    normalize = "panels"
  ) +
  geom_text(
    aes(
      label = glue::glue("p = {round(p, 2)}")
    ),
    x = 0,
    hjust = 1.1,
    position = position_nudge(x = -.1, y = .12),
  ) +
  geom_vline(xintercept = 0, color = "grey20", lty = 2) +
  scale_fill_manual(
    values = c("grey75", "skyblue2"),
    name = "Effect Direction",
    breaks = c(FALSE, TRUE),
    labels = c(
      "Negative",
      "Positive"
    ),
    guide = guide_legend()
  ) +
  labs(
    x = "Magnitude",
    y = "Likelihood"
  ) +
  bptheme::theme_blueprint(grid = "Xxy", grid_col = "grey80") +
  theme(strip.placement = "outside")
```

# Employment, education, and social assistance receipt 

```{r binary-var}
d_emp <- tibble::tribble(
  ~outcome, ~coefficient, ~estimate, ~str_error, ~t, ~p,
  "Employment", "Intercept", 0.27143, 0.05622, 4.828, 3.44e-06,
  "Employment", "Treatment", 0.11319, 0.07744, 1.462, 0.146,
  "Enrolled in Education", "Intercept", 0.21429, 0.04884, 4.388, 2.18e-05,
  "Enrolled in Education", "Treatment", -0.01175, 0.06707, -0.175, 0.861,
  "Social Assistance Receipt", "Intercept", 0.54286, 0.06000, 9.048, 8.03e-16,
  "Social Assistance Receipt", "Treatment", -0.01121, 0.08240, -0.136, 0.892
)

d_emp_plt <- d_emp %>%
  select(outcome, coefficient, estimate) %>%
  pivot_wider(names_from = coefficient, values_from = estimate) %>%
  rename(
    treatment_effect = `Treatment`,
    control = `Intercept`
  ) %>%
  mutate(
    treatment = treatment_effect + control,
    scale_range = case_when(
      outcome == "Employment" ~ 1 - 0,
      outcome == "Enrolled in Education" ~ 1 - 0,
      outcome == "Social Assistance Receipt" ~ 1 - 0
    )
  ) %>%
  pivot_longer(
    cols = c(control, treatment),
    names_to = "group",
    values_to = "Scale units"
  ) %>%
  mutate(
    `% of scale range` = `Scale units` / scale_range * 100
  ) %>%
  pivot_longer(
    cols = c(`Scale units`, `% of scale range`),
    names_to = "transformation",
    values_to = "observed_change"
  ) %>%
  mutate(
    transformation = fct_inorder(transformation)
  )

d_emp_plt %>%
  ggplot(aes(y = group, x = observed_change)) +
  facet_grid(rows = vars(outcome), cols = vars(transformation), scales = "free_x", switch = "y") +
  geom_segment(aes(xend = 0, yend = after_stat(y))) +
  geom_point(shape = 21, size = 3, fill = "#0000FF") +
  geom_text(
    aes(label = round(observed_change, 1)),
    position = position_nudge(y = .12)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(.1, .1))
  ) +
  labs(
    x = "Average Change in Score",
    y = NULL
  ) +
  bptheme::theme_blueprint(grid = "Xxy", grid_col = "grey80") +
  theme(strip.placement = "outside")

## Generalized lienar model

job_employed <- dat_rct_prep$job_employed %>%
  select(enrollment, job_employed, survey_name, experiment_assignment_group) %>%
  mutate(
    survey = str_extract(survey_name, "Baseline|Program Exit|3 Months"),
    survey = factor(survey)
  ) %>%
  select(-survey_name) %>%
  mutate(experiment_assignment_group = factor(experiment_assignment_group)) %>%
  mutate(job_emp = case_when(
    job_employed == "Yes (employee and/or self-employed)" ~ 1,
    job_employed == "No" ~ 0,
    TRUE ~ NA
  )) %>%
  select(-job_employed) %>%
  pivot_wider(
    names_from = survey,
    values_from = job_emp
  ) %>%
  pivot_longer(
    cols = -c(enrollment, experiment_assignment_group),
    names_to = "survey",
    values_to = "job_emp"
  ) %>%
  pivot_wider(
    names_from = survey,
    values_from = job_emp
  ) %>%
  filter(!is.na(Baseline) & !is.na(`Program Exit`) & !is.na(`3 Months`)) %>%
  mutate(assignment = case_when(
    experiment_assignment_group == "IMM+ Treatment group" ~ TRUE,
    experiment_assignment_group == "IMM+ Comparison group" ~ FALSE
  )) %>%
  rename(enrollment_id = enrollment) %>%
  select(-experiment_assignment_group)

# Conduct a logistic regression on the binary variable

emp_glm <- glm(
  formula = `3 Months` ~ assignment,
  data = job_employed,
  family = "binomial"
)

# Print coefficients from logistic regression. These are logits. We can see effect of treatment on employment status at 3 month is positive. Being in treatment group increases the probability of being employed at 3 month.

summary(emp_glm)$coefficients

# source logit to probability conversion funtion

source("https://sebastiansauer.github.io/Rcode/logit2prob.R")

logit2prob(coef(emp_glm))

# interpreting results:
# the employment probability is 0.2714286 if in control group
# logistic regression can be interpreted as a normal regression as long as you use logits

# logits_employed = intercept + b_employed*assignment (b_employed is given in logits)

predict(emp_glm, data.frame(assignment = TRUE), type = "response")

predict(emp_glm, data.frame(assignment = FALSE), type = "response")
```

```{r}
grid <- tibble::tibble(
  assignment = c(TRUE, FALSE)
)

pred <- predict(emp_glm, grid, type = "link", se.fit = TRUE)

grid <- grid |>
  mutate(
    link = pred$fit,
    se = pred$se.fit,
    sim_link = map2(link, se, rnorm, n = 10000)
  ) |>
  unnest(sim_link) |>
  mutate(
    sim_prob = exp(sim_link) / (1 + exp(sim_link))
  )

grid |>
  ggplot(aes(x = sim_prob, y = assignment)) +
  stat_halfeye() +
  scale_x_continuous(limits = c(0, 1))
```

